Different forms of uncertainty differently affect the evolution of social learning

### 2022-08-13 ms.tex
finished revisions with N=1000 and full supplement

@@

Many models of social learning assume little to no cognitive capabilities
of their agents. For example, a common modelling strategy compare the payoffs of
agents with different pure learning strategies (some will always engage in social
learning, and others in individual learning) while revealing nothing about the
cognition underlying individuals' decisions  ~\cite{BoydRicherson1985, Rogers1988,
aoki2005}. Some more complicated learning strategies that make use of both
socially and individually learned information have been proposed
~\cite{Enquist2007, perreault2012bayesian}, but these are seldom designed to
reflect cognitive mechanisms for integrating information. Though such
simplifications are often useful, several authors have warned of the risks to
\emph{blackboxing}, or ignoring, cognitive mechanisms \cite[p. 658]{Heyes2016,
Kendal2018}, given such details may be theoretically meaningful. In order to
understand how social learning evolves as an adaptation to uncertainty, we need to
endow agents with a biologically plausible cognition capable of integrating
information from various sources in a payoff-maximizing way~\cite{Gershman2019}. 
% \cm{I don't
%   understand this next sentence, but think it can be cut with no loss of meaning:
%   In our model we do not impose any structure in randomness between behaviors, so
% our cognitive model is only sensitive to total randomness.} 
To do so, we let
agents see the payoffs to behaviors and use the softmax function to convert these
observations to a probability distribution from which they strategically select
their own behavior. %\cm{does this still make sense? and does it need a citation?}

@@

To understand how uncertainty affects social learning, 
we developed an agent-based model of
social learning evolution.  Evolutionary agent-based models simulate changes in
prevalence of a phenotype over several generations---in this model the phenotype
is whether or not an agent engages in social learning.  In the model, a
population of $N$ individuals each must decide which of $B$ behaviors to perform at
each time step within a generation consisting of $L$ time steps. At the end of each
generation, agents are selected with replacement to reproduce another $N$ agents, 
biased by net payoffs. Agents
who inherit the social learning trait then learn about payoffs from the
previous generation, while asocial learners begin life with no \emph{a priori}
knowledge of the world.  Each behavior is a ``bandit'', a common modeling and
experimental approach for representing behaviors with Bernoulli-distributed
payoffs~\cite{SuttonBartoBook,McElreath2005,Steyvers2009,Rendell2010,Schulz2019}.
All bandits thus yield a payoff of 1 or 0 with a certain probability.  One of the
bandits/behaviors in the model is more likely to pay off than all the rest.  Agents
optimize their net payoffs over their lifespan by quickly figuring out which
behavior is optimal and performing it as often as possible within their lifespan.

We operationalized four separable varieties of uncertainty that affect the
ability of agents to find and choose the optimal behavior: 
\emph{environmental variability}; 
\emph{selection set size}; \emph{payoff ambiguity}; and the \emph{effective lifespan}. 
Environmental variability is the probability that the optimal behavior changes
between generations. We do not consider within-generation variability. 
Furthermore, our model is not spatial, so agents have free, instant
access to all behaviors, and there is no spatial uncertainty. 
These conditions reflect humans in a computer-based social learning evolution
experiment~\cite{McElreath2005,Morgan2012,Derex2016}. Spatial dimensions could also be
considered irrelevant if costs associated with accessing a behavior were equal,
e.g., if it was assumed to be equally costly for animals to
visit alternative food sources~\cite{Aplin2017,Baracchi2018}.  
% This is empirically justified since cultural evolution
% occurs when meta-cognitive strategies and information are transferred via
% learning and social influence, instead of by physical reproduction and death
% as in genetic evolution.  
The selection set size is the total number of possible
behaviors agents choose from, assumed to be constant within a simulation.  
Payoff ambiguity is the payoff difference between optimal
and non-optimal behaviors. The effective life span is the
number of time steps in which agents perform behaviors within a generation. 
Each simulated environment provides a distinct challenge
to the agents, which will lead agents to differently evolve to become social
learners or not, depending on which strategy optimizes net payoffs.

\emph{Agents} here are autonomous, intelligent, simulated problem solvers. Each
agent evaluates its environment and chooses how to act within that environment based
on a set of empirically-motivated rules.  Our agents select which behavior to do
probabilistically, weighted by the softmax function applied to the mean payoff they
have observed for each behavior. The softmax function is a biologically plausible
generalization of behavior selection under uncertainty that enables agents to often
greedily exploit the most lucrative behavior (based on their observations) but also
to sometimes explore alternatives~\cite{Schulz2019,Collins2013,Daw2006,Yechiam2005}.
Such individual-level intelligence is a critical adaptation that has 
co-evolved with the capacity for social learning, and we will show this has a marked influence 
on the cultural evolution of social learning.

In \emph{evolutionary} agent-based models like ours, agent life cycles are
programmed to cause agents to reproduce and die off, where reproducers pass on
heritable traits to their offspring. 
Natural selection occurs in the model when agents with higher
payoffs preferentially reproduce. This is a model of cultural evolution, so
``reproduction'' means teaching the meta-cognitive strategy of whether one should learn
socially at all; ``death'', in this interpretation, means
loss of the potential to socially influence others. 
This sets up an inter-generational structure
that we also use as a scaffold for \emph{oblique} social learning, meaning children
with the social learning trait learn about payoffs from a member of the
previous generation. Evolution acts as an optimization routine to identify
whether individuals with the social learning trait will outperform those without.

Agents inherit one trait in this model, whether or not
an agent, $i$, learns socially. If $i$ has the trait we write $s_i=1$
 and 0 otherwise. In our analysis, we 
average $s_i$ over all agents and simulation trials to understand the affects of
different uncertainty parameter settings. 
Our computational analysis amounts to observing patterns in
the aggregated $s_i$, supported by measuring the number of generations 
to model convergence and expected payoffs. 
To understand how different varieties of uncertainty interact
to affect the evolution of social learning, we systematically created an array of
simulated environments by systematically varying the uncertainty parameters.

@@

We performed sensitivity analyses to ensure that our main analysis is reasonably
robust to a range of auxiliary parameters. First,
we varied the population size, $N$, to explore whether smaller $N$ would 
show more drift due to finite population size effects, but also generally support
our main conclusions. Next, we varied the
softmax greediness parameter $\beta \in \{1, 100\}$ to supplement the main results
that used $\beta = 10$. We do not
expect agents with different $\beta$ to perform equally well 
individually, so certain $\beta$ values may themselves suppress social learning
by undermining individual learning. Nonetheless, our results
should be valid over some range of $\beta$. Finally, we varied the
number of prospective teachers $N_T \in \{2, 10, 20\}$ 
to supplement the main results that used $N_T = 5$. The number of prospective
teachers should not change whether social learning is optimal, 
but it may induce more drift since the benefit of social learning may
be more difficult to detect with fewer prosopective teachers.

@@

To understand how different forms of uncertainty interact to affect the evolution of
social learning, we measured how frequently social learning evolved across
uncertainty parameters. We analyzed outcomes by plotting the mean observed social
learning frequency, $\meansl$, with mean taken across all agents and 1000 simulation trials for each
uncertainty parameter setting.   
% This is convenient because we are most certain about how $\meansl$ depends on $u$:
% we expect $\meansl$ to decrease as $u$ increases since environmental variability
% is well understood to cause socially-transmitted information to become outdated. 
% To inspect patterns in $\meansl$ over all uncertainty parameters, 
We analyze nine plots of how $\meansl$ (y-axis in each plot) changes over $u$ (x-axis) 
in a $3\times3$ plot array (Figure~\ref{fig:mainResults}). 
% All plots show how $\meansl$ changes (y-axis)
% with environmental variability, $u$ (x-axis). 
Selection set size, $B$ increases from left to right in the plot array (columns). 
Payoff ambiguity increases as $\pilow$ increases from top to bottom in the array 
(recall $\pihigh=0.9$ for all analyses).
Effective lifespan, $L$, is indicated by line and marker colors (inset). 
Several patterns in social learning prevalence emerged as 
these uncertainty parameters were varied (Figure~\ref{fig:mainResults}). 
% We support our conclusions regarding the evolution of social learning by analyzing
% the time to fixation and by comparing the average net payoffs in our simulations
% to the expected payoffs for individuals in both populations of all social learners
% and in populations of all individual learners. We thus use our analysis to
% understand changes to how $\meansl$ changes over $u$ for different uncertainty
% parameter settings.

First, we confirmed existing theoretical predictions that greater $u$
uniformly decreased $\meansl$ across all other uncertainty parameter values.
The other uncertainty parameters changed the $u$ value at which $\meansl$
began to decrease, and how rapidly $\meansl$ decreased over $u$.  For example,
when payoff ambiguity and selection set size were both small ($\pilow = 0.1$
and $B=2$), social learning evolved only when $u$ was small. 
% These parameter settings present a relatively easy problem for individuals.
% Furthermore, it is relatively costly for agents to be misled by outdated
% social information in this case, especially for longer lifespans for finding
% the optimal behavior (Figure~\ref{fig:mainResults}, upper left). 
When $\pilow = 0.8$ and $B=2$, for exapmle, $\meansl$ decreased less rapdily
over $u$ due to drift caused by weak selection caused by little advantage for
performing the optimal versus non-optimal behavior (see rows of
Figure~\ref{fig:mainResults}). We analyzed several such patterns of social
learning evolution and suppression across different uncertainty settings.

Increased selection set size $B$ led nearly uniformly to more prevalent social
learning, with all other uncertainty parameters held constant. This is because,
first, it is more difficuilt to find the optimal behavior with a larger $B$.
% more options present a more difficult problem for trial-and-error learning within
% a lifespan. 
Second, this effect is amplified because agents in a world with larger $B$ are more
cognitively flexible compared to those in a smaller $B$ world. 
When $B$ is larger, agents are relatively more likely to choose an alternative to
their current known optimum. %is easier to correct for larger $B$ relative to smaller $B$.  
To understand this, consider the case where a social learner child learns
that behavior 1 has an expected payoff of 0.8, and the rest have an expected
payoff of 0, but the optimal behavior changed between generations. Assume
the child tries behavior 1, but it pays off 0.
If $B=2$, then at the next within-generation time step, $\ledger = (0.4, 0.0)$. The probability
the agent explores behavior 2 is now 0.02. However, if $B=10$ the agent has
$\ledger = (0.4, 0.0, \ldots, 0.0)$; the probability
the agent explores one of the nine alternative behaviors increases to 0.14.
If the agent again tries behavior 1 and it pays 0, the probability of exploring alternatives
with $B=2$ is 0.21, but with $B=10$ the probability of exploring alternatives
greatly increases to 0.71. \mt{I double checked these calculations, but would appreciate if
someone triple checked it}. This explanation does not apply when $L=1$, where
individual learning does not occur.

Payoff ambiguity had a modest effect when $\pilow = 0.1$ increased to $\pilow =
0.45$, causing social learning to be favored across more values of $u$ (first two rows of
Figure~\ref{fig:mainResults}).  This is because increased payoff ambiguity increases
the difficulty, and reduces the overall benefit, of discerning which behavior is
optimal.  When $\pilow = 0.8$ the
$\meansl$ curves become smoother, indicating more evolutionary drift towards 
possibly non-optimal fixations (bottom row of
Figure~\ref{fig:mainResults}). For smaller values of $u$, this results in less
frequent evolution of social learning for the same values of $u$ and $B$. But past
the value of $u$ where $\meansl \approx 0.5$ this switches, and social learning evolves more
frequently than for smaller $\pilow$. 
% When $\pilow \to \pihigh$, agents 
% accumulate increasingly more payoffs, whether they are social learners or not. 
Greater payoff ambiguity means lower risk to try possibly outdated social
information by reducing the penalty for performing non-optimal behaviors.

We observed two distinct trends in the evolution of social learning as
$L$ was varied.  First, when $\pilow$ was sufficiently small ($\pilow=0.1, 0.45$),
we observed that shorter lifespan extended the evolution of social learning over more values of
$u$, since payoff-biased social learning means teachers are reliable.
When $\pilow=0.8$, shorter lifespans decreased $\meansl$ in some cases and
increased $\meansl$ in other cases due to greater evolutionary drift. In
this setting, many agents will obtain a payoff of 1 at each time step in their 
lifetime, which, combined with finite population size, will make it more likely for 
evolution to settle on a non-optimal population fixation.
% In the first
% case, $\meansl$ was reduced for $L=1$ for small values of $u$, when other $L > 1$
% led to $\meansl \approx 1$. At larger $u$ , the $L=1$ case yields non-zero
% $\meansl$, while for the other $L > 1$, $\meansl \approx 0$. 
Smaller $L$ means agents end their lives more uncertain about which behavior was
optimal, but payoff-biased teacher selection and learning can remedy this if
payoff ambiguity is sufficiently small.
% When $L$ is smaller, agents have fewer opportunities to learn through trial and
% error, so it is more difficult for them to discern which strategy is optimal, social
% or asocial.

@@

To explain the evolution of social learning we assumed that, first, the relative benefit
of having the social learning trait is essential to its evolution; and, second,
we assumed that values of $\meansl \neq 0,1$ indicate evolutionary drift---but does our
data support these assumptions?  Indeed, social learning evolution and evolutionary
drift depended on the relative benefit of the social learning trait. We measured
drift in terms of the average number of generations to fixation, denoted $\langle G
\rangle$. Indeed, social learning evolution and drift are reflected in the relative
benefit of social learning to asocial learning.  When social and asocial learning are
equally beneficial, fixation takes longer to achieve, indicating greater drift.

First, we confirmed that social learning evolves when $\meansoc > \meanasoc$.
% by first comparing the expected
% payoff for populations initialized as either all individual learners or all social
% learners, denoted $\meansoc$ and $\meanasoc$, respectively.  
In the case of $L=1$ we can calculate $\meanasoc$ directly: it is just the
expected payoff across all bandits. When $L=8$, and for all uncertainty parameters 
for $\meansoc$, we calculate expected social and asocial payoffs by respectively setting
$\sum_i = N$ or $\sum_i = 0$ at model initialization. These stay constant since
we do not allow for mutation.  We show expected payoffs for these four settings in this main text; we show the same
plots for $\pilow = 0.8$ in the supplement both for $N=1000$ and $N=100$.

Further inspection of maximal payoff ambiguity ($\pilow = 0.8$) showed strong
effects of population size under sensitivity analysis\mt{TODO; not currently included in 
the appendix}. In the default setting
of $N=1000$, populations of initially all social learners performed much worse
in many uncertainty contexts than either the actual payoffs accumulated by 
model agents, and lower than even the expected individual-level payoffs,
even though social learning evolved in many of the same contexts in our model
with the default settings (Figure SupplementX1). When the population was
reduced to $N=100$, there was a tighter evolutionary bottleneck on 
underperforming agents, and selection more effectively weeded out underperforming agents.
This demonstrates that larger populations not only can support beneficial
choices~\cite{Henrich2004}, but also maladaptive information.

@@

We confirmed that drift increased when $\meansl$ transitioned from 1 to 0, 
and when $\pilow$ was increased. We measured drift by the average number of 
generations to fixation, $\meanG$. Recall fixation is defined as the population
becoming homogenous, composed of all social or all asocial learners. 
$\meanG$ peaks around the values of $u$ 
when $\meansoc \approx \meanasoc$, which led $\meansl$ to transition from 1 to 0.
In some cases, especially when $\pilow=0.8$,
there is no distinct peak in $\meanG$ across $B$ and $L$ settings, 
indicating little difference in task difficulty across a range of uncertainty parameter settings.

@@

By disambiguating and organizing various operationalizations of uncertainty
into a single computational model, we developed a more nuanced theoretical
understanding of how different forms of uncertainty interact to affect the
evolution of social learning.  We reviewed and distilled several empirical
and theoretical models in the literature along four uncertainty dimensions:
temporal environmental uncertainty, selection set size, ambiguous payoffs, and
effective lifespan.  The model reproduced classic predictions that social
learning can yield higher payoffs than individual learning when uncertainty is
sufficiently limited. Our model also demonstrated a new insight into social
learning as a scaffold for individual learning, enabled by endowing agents
with an evolutionarily- and biologically-plausible individual learning
mechanism, softmax search. Specifically, we found that agents could recover
from receiving outdated information, so social learning could evolve even
though social information was outdated more often than not. We found increased
payoff ambiguity and lifespan led to greater drift, which is essentially
evolutionary uncertainty in a finite-size population. Detailed theoretical
models and predictions of cultural evolution under uncertainty are critical
for understanding how humans adapt to an uncertain and rapidly changing world,
both to mitigate existential threats~\cite{Moya2020,Jones2021}, especially in
the most vulnerable communities~\cite{McNamara2020}, and to capitalize on new
behavioral opportunities such as transitioning to clean energy use and
production~\cite{NatureEnergyEditorialPromisesPremises2018,Brisbois2022}.

@@

We found that the evolution of social learning, under a few modest assumptions, is a
complex phenomenon, dependent on several interrelated factors. By systematically
identifying and operationalizing common theoretical assumptions about contextual
uncertainty in social learning models, we were able to develop a more systematic
understanding of which uncertainty factors interact to determine whether social learning
evolves. By explicitly modeling a key individual-level learning mechanism shared across taxa, namely the
ability to adjust behavior to uncertainty and new information, we saw that social
learning could evolve even if social learning regularly provides outdated
information. This work informs most directly our current understanding of the evolution of
social learning, but has broader significance for understanding problem
solving under uncertainty. Our model and its software implementation were
designed to be modular and extensible.  Indeed, we plan to extend this model to further
develop a more detailed theoretical understanding of the evolution of social learning;
we hope others will, too.

@@

Model populations reached fixation between few to 140 generations on
average (Figure~\ref{fig:steps}).
Note 140 generations is equivalent to 2800 time steps. We set the maximum 
number of iterations globally to $20~000$, which is 1000 generations when
$L=20$. Although we can be relatively sure that this maximum number of 
generations is rarely reached based on our main text analysis of $\meanG$,
we additionally tabulated the number of simulations that did not converge to
fixation
within this step limit. We found that 0.1\% of simulations did not converge
when $B=2$, 0.2\% did not converge when $B=4$, and 2.0\% did not converge
when $B=10$ (Table~\ref{tab:convergence}). When a trial did not converge we
still included that trial's $\meansl$ value in the average across trials,
but for these $0 < \meansl < 1$.

@@

\newpage
\subsection{Number of prospective teachers sensitivty analysis}

Slightly more drift when $N_T = 2$, slightly less drift when $N_T = 20$, 
but mostly just reproduces results for main text setting $N_T = 5$.

>>>>

\usepackage{xcolor}
% \renewcommand{\linenumberfont}{\small\color{gray}}

@@

Many models of social learning assume little to no cognitive capabilities
of their agents. For example, a common modelling strategy compares the payoffs of
agents with different pure learning strategies (some will always engage in social
learning, and others in individual learning) while revealing nothing about the
cognition underlying individuals' decisions~\cite{BoydRicherson1985, Rogers1988,
aoki2005}. Some more complicated learning strategies that make use of both
socially and individually learned information have been proposed
~\cite{Enquist2007, perreault2012bayesian}, but these are seldom designed to
reflect cognitive mechanisms for integrating information. Though such
simplifications are often useful, several authors have warned of the risks to
\emph{blackboxing}, or ignoring, cognitive mechanisms \cite[p. 658]{Heyes2016,
Kendal2018}, given such details may be theoretically meaningful. In order to
understand how social learning evolves as an adaptation to uncertainty, we 
endowed agents with a biologically plausible cognition capable of integrating
information from various sources to maximize payoffs, \emph{softmax
search}~\cite{Gershman2019}, described in more detail below. 
% \cm{I don't
%   understand this next sentence, but think it can be cut with no loss of meaning:
%   In our model we do not impose any structure in randomness between behaviors, so
% our cognitive model is only sensitive to total randomness.} 
% To do so, we let
% agents see the payoffs to behaviors and use the softmax function to convert these
% observations to a probability distribution from which they strategically select
% their own behavior. %\cm{does this still make sense? and does it need a citation?}

@@

To understand how uncertainty affects social learning, 
we developed an agent-based model of
social learning evolution.  Evolutionary agent-based models simulate changes in
the prevalence of a phenotype over several generations---in this model the phenotype
is whether or not an agent engages in social learning.  In our model, a
population of $N$ individuals each must decide which of $B$ behaviors to perform at
each time step within a generation consisting of $L$ time steps. At the end of each
generation, agents are selected with replacement to reproduce another $N$ agents, 
biased by net payoffs. Agents
who inherit the social learning trait then learn about payoffs from the
previous generation, while asocial learners begin life with no \emph{a priori}
knowledge of the world.  Each behavior is a ``bandit'', a common modeling and
experimental approach for representing behaviors with random-valued
payoffs~\cite{SuttonBartoBook,McElreath2005,Steyvers2009,Rendell2010,Schulz2019}.
Bandits in our model yield Bernoulli payoffs of 1 or 0 with some probability,
with that probability defining each behavior's expected payoffs.  One of the
bandits/behaviors in the model is more likely to pay off than all the rest.  Agents
optimize their net payoffs over their lifespan by quickly figuring out which
behavior is optimal and performing it as often as possible within their lifespan.
The optimal behavior is static within agent lifespans, but varies in general between
generations, representing environmental variability.

We operationalized four separable varieties of uncertainty that affect the
ability of agents to find and choose the optimal behavior: 
\emph{environmental variability}; 
\emph{selection set size}; \emph{payoff ambiguity}; and the \emph{effective lifespan}. 
Environmental variability is the probability that the optimal behavior changes
between generations.   
% This is empirically justified since cultural evolution
% occurs when meta-cognitive strategies and information are transferred via
% learning and social influence, instead of by physical reproduction and death
% as in genetic evolution.  
The selection set size is the total number of possible
behaviors agents choose from, assumed to be constant within a simulation.  
Payoff ambiguity is the payoff difference between optimal
and non-optimal behaviors. The effective life span is the
number of time steps in which agents perform behaviors within a generation. 
Our model is not spatial, so agents have free, instant
access to all behaviors, and there is no spatial uncertainty. 
These conditions reflect humans in a computer-based social learning evolution
experiment~\cite{McElreath2005,Morgan2012,Derex2016}. Spatial dimensions could also be
considered irrelevant if costs associated with accessing a behavior were equal,
e.g., if it was assumed to be equally costly for animals to
visit alternative food sources~\cite{Aplin2017,Baracchi2018}.
Each simulated environment provides a distinct challenge
to the agents, which will lead agents to differently evolve to become social
learners or not, depending on which strategy optimizes net payoffs.

\emph{Agents} here are autonomous, intelligent, simulated problem solvers. Each
agent evaluates its environment and chooses how to act within that environment based
on a set of empirically-motivated rules.  Our agents select which behavior to do
probabilistically, weighted by the softmax function applied to the mean payoff they
have observed for each behavior. The softmax function is a biologically plausible
generalization of behavior selection under uncertainty that enables agents to often
greedily exploit the most lucrative behavior (based on their observations) but also
to sometimes explore alternatives~\cite{Schulz2019,Collins2013,Daw2006,Yechiam2005}.
Such individual-level intelligence is a critical adaptation that has 
co-evolved with the capacity for social learning. We will show that individual
learning can support the evolution of social learning.

In our \emph{evolutionary} agent-based model, agent life cycles are
programmed to cause agents to reproduce and die off, where reproducers pass on
heritable traits to their offspring. 
Natural selection occurs in the model when agents with higher
payoffs preferentially reproduce. This is a model of cultural evolution, so
``reproduction'' means teaching the meta-cognitive strategy of whether one should learn
socially at all; ``death'', in this interpretation, means
loss of the potential to socially influence others. 
This sets up an inter-generational cultural structure
that we also use as a scaffold for \emph{oblique} social learning, meaning children
with the social learning trait learn about payoffs from a member of the
previous generation. Evolution acts as an optimization routine to identify
whether individuals with the social learning trait will outperform those without.

Agents inherit one trait in this model, whether or not
an agent, $i$, learns socially. If $i$ has the trait we write $s_i=1$
 and 0 otherwise. In our analysis, we 
average $s_i$ over all agents and simulation trials to understand the affects of
different uncertainty parameter settings. 
Our computational analysis is based on observing and comparing patterns in
aggregated $s_i$ across different parameter settings, 
supported by measuring the number of generations 
to model convergence and expected payoffs in populations across the same contexts. 
To understand how different varieties of uncertainty interact
to affect the evolution of social learning, we systematically created an array of
simulated environments by systematically varying the uncertainty parameters, and
auxiliary model parameters to ensure model validity.

@@

We performed sensitivity analyses to ensure that our main analysis is reasonably
robust to a range of auxiliary parameters. First,
we varied the population size, $N$, to explore whether smaller $N$ would 
show more drift due to finite population size effects, but also generally support
our main conclusions. Next, we varied the
number of prospective teachers, $N_T \in \{2, 10, 20\}$, 
to supplement the main results that used $N_T = 5$. The number of prospective
teachers should not change whether social learning is optimal, 
but it may induce more drift since the benefit of social learning may
be more difficult to detect with fewer prosopective teachers. Finally, we varied the
softmax greediness parameter $\beta \in \{1, 100\}$ to supplement the main results
that used $\beta = 10$. We do not
expect agents with different $\beta$ to perform equally well 
individually, so certain $\beta$ values may themselves suppress social learning
by undermining individual learning. Nonetheless, our results
should be valid over some range of $\beta$.

@@

To understand how different forms of uncertainty interact to affect the evolution of
social learning, we measured how frequently social learning evolved across
uncertainty parameters. We analyzed outcomes by plotting the mean observed social
learning frequency, $\meansl$, with mean taken across all 1000 agents and 1000
simulation trials for each uncertainty parameter setting at model fixation.
% This is convenient because we are most certain about how $\meansl$ depends on $u$:
% we expect $\meansl$ to decrease as $u$ increases since environmental variability
% is well understood to cause socially-transmitted information to become outdated. 
% To inspect patterns in $\meansl$ over all uncertainty parameters, 
We analyze nine plots of how $\meansl$ (y-axis in each plot) changes over 
environmental variability, $u$ (x-axis),
in a $3\times3$ plot array (Figure~\ref{fig:mainResults}). 
% All plots show how $\meansl$ changes (y-axis)
% with environmental variability, $u$ (x-axis). 
Selection set size, $B$, increases from left to right in the plot array (columns). 
Payoff ambiguity increases as the non-optimal payoff, $\pilow$, increases from 
top to bottom in the array (recall $\pihigh=0.9$ for all analyses).
Effective lifespan, $L$, is indicated by line and marker colors (inset). 
Several patterns in social learning prevalence emerged as 
these uncertainty parameters were varied (Figure~\ref{fig:mainResults}).

First, we confirmed existing theoretical predictions that increased $u$
monotonically decreased $\meansl$ across all other uncertainty parameter values.
The other uncertainty parameters changed the $u$ value at which $\meansl$
began to decrease, and how rapidly $\meansl$ decreased over $u$.  For example,
when payoff ambiguity and selection set size were both small ($\pilow = 0.1$
and $B=2$), social learning evolved only when $u$ was small. 
% These parameter settings present a relatively easy problem for individuals.
% Furthermore, it is relatively costly for agents to be misled by outdated
% social information in this case, especially for longer lifespans for finding
% the optimal behavior (Figure~\ref{fig:mainResults}, upper left). 
When $\pilow = 0.8$ and $B=2$, for exapmle, $\meansl$ decreased less rapdily
over $u$ due to drift caused by weak selection due to little advantage for
performing the optimal versus non-optimal behavior (see rows of
Figure~\ref{fig:mainResults}). We analyzed several such patterns of social
learning evolution and suppression across different uncertainty settings.

Increased selection set size $B$ led nearly uniformly to more prevalent social
learning, with all other uncertainty parameters held constant. This is because, first,
it is more difficuilt to find the optimal behavior with a larger $B$.  Second, this
effect is amplified because agents in a world with larger $B$ are more cognitively
flexible compared to those in a smaller $B$ world.  When $B$ is larger, agents are
relatively more likely to choose an alternative to their current known optimum.  To
understand this, consider the case where a social learner child learns that behavior 1
has an expected payoff of 0.8, and the rest have an expected payoff of 0, but the
optimal behavior changes between generations. Assume the child tries behavior 1, but
it pays off 0.  If $B=2$, then at the next time step, $\ledger =
(0.4, 0.0)$. The probability the agent explores behavior 2 is now 0.02. However, if
$B=10$ the agent has $\ledger = (0.4, 0.0, \ldots, 0.0)$, and the probability the agent
explores one of the nine alternative behaviors increases to 0.14.  If the agent again
tries behavior 1 and it pays 0, the probability of exploring alternatives with $B=2$
is 0.21, but with $B=10$ the probability of exploring alternatives more greatly increases
to 0.71. \mt{I double checked these calculations, but would appreciate if someone
  triple checked it}. 
This explanation does not apply when $L=1$, where individual learning does not occur.

Payoff ambiguity had a modest effect when $\pilow = 0.1$ increased to $\pilow =
0.45$, causing social learning to be favored across more values of $u$ (first two rows of
Figure~\ref{fig:mainResults}).  This is because increased payoff ambiguity increases
the difficulty, and reduces the overall benefit, of discerning which behavior is
optimal.  When $\pilow = 0.8$ the
$\meansl$ curves become smoother, indicating more evolutionary drift towards 
possibly non-optimal fixations (bottom row of
Figure~\ref{fig:mainResults}). For smaller values of $u$, this results in less
frequent evolution of social learning for the same values of $u$ and $B$. But past
the value of $u$ where $\meansl \approx 0.5$ this switches, and social learning evolves more
frequently than for smaller $\pilow$. 
% When $\pilow \to \pihigh$, agents 
% accumulate increasingly more payoffs, whether they are social learners or not. 
Greater payoff ambiguity means lower risk to try possibly outdated social
information by reducing the penalty for performing non-optimal behaviors, leading
to the observed drift in aggregated results.

We observed two distinct trends in the evolution of social learning as
$L$ was varied.  First, when $\pilow$ was sufficiently small ($\pilow=0.1, 0.45$),
we observed that shorter lifespan extended the evolution of social learning over more values of
$u$, since payoff-biased social learning means teachers are reliable.
When $\pilow=0.8$, shorter lifespans decreased $\meansl$ in some cases and
increased $\meansl$ in other cases due to greater evolutionary drift. In
this setting, many agents will obtain a payoff of 1 at each time step in their 
lifetime, which, combined with finite population size, will make it more likely for 
evolution to guide the population to non-optimal social/asocial learning stragegy.
% In the first
% case, $\meansl$ was reduced for $L=1$ for small values of $u$, when other $L > 1$
% led to $\meansl \approx 1$. At larger $u$ , the $L=1$ case yields non-zero
% $\meansl$, while for the other $L > 1$, $\meansl \approx 0$. 
Smaller $L$ means agents end their lives more uncertain about which behavior was
optimal, but payoff-biased teacher selection and learning can remedy this if
payoff ambiguity is sufficiently small.
% When $L$ is smaller, agents have fewer opportunities to learn through trial and
% error, so it is more difficult for them to discern which strategy is optimal, social
% or asocial.

@@

To explain the evolution of social learning we implicitly assumed above that, first,
the relative benefit of having the social learning trait is essential to its
evolution; and, second, we assumed that values of $\meansl \neq 0,1$ indicate
evolutionary drift---but do our data support these assumptions?  Indeed, we found that
social learning evolution and evolutionary drift depended on the relative benefit of
the social learning trait. We measured drift in terms of the average number of
generations to fixation, denoted $\langle G \rangle$. Indeed, social learning
evolution and drift are reflected in the relative benefit of social learning to
asocial learning.  When social and asocial learning are equally beneficial, fixation
takes longer to achieve, indicating greater drift.

First, we confirmed that social learning evolves when the expected payoffs to social
learners exceed those of asocial learners, which we write $\meansoc > \meanasoc$.
In the case of $L=1$ we can calculate $\meanasoc$ directly: it is just the expected
payoff across all bandits. When $L=8$, and for all uncertainty parameters for
$\meansoc$, we calculate expected social and asocial payoffs by respectively setting
$\sum_i = N$ or $\sum_i = 0$ at model initialization. These stay constant since we do
not allow for mutation.  
In general, we observed that social learning evolved when its expected payoffs
(Figure~\ref{fig:payoffs}, diamond-dash lines) exceeded expected individual-learner
payoffs (Figure~\ref{fig:payoffs}, long-dashed lines). 
However, we also found that populations of initially all social learners performed worse
in many uncertainty contexts than either the actual payoffs accumulated by 
model agents, and lower than even the expected individual-level payoffs,
even though social learning evolved in many of the same contexts in our model,
especially when $\pilow = 0.8$ 
(Figure~\ref{fig:fullMeanPrevNetPayoffs}, bottom row).

@@

We confirmed that drift increased when $\meansl$ transitioned from 1 to 0, 
and when $\pilow$ was increased (Figure~\ref{fig:steps}). 
We measured drift by the average number of 
generations to fixation, $\meanG$. Recall fixation is defined as the population
becoming homogenous, composed of all social or all asocial learners. 
$\meanG$ peaks around the values of $u$ 
when $\meansoc \approx \meanasoc$, which led $\meansl$ to transition from 1 to 0.
In some cases, especially when $\pilow=0.8$,
there is no distinct peak in $\meanG$ across $B$ and $L$ settings, 
indicating little difference in task difficulty across a range of uncertainty
parameter settings (Figure~\ref{fig:steps}, bottom row).

@@

By disambiguating and organizing various operationalizations of uncertainty
into a single computational model, we developed a more nuanced theoretical
understanding of how different forms of uncertainty interact to affect the
evolution of social learning.  We reviewed and distilled several empirical
and theoretical models in the literature along four uncertainty dimensions:
temporal environmental uncertainty, selection set size, ambiguous payoffs, and
effective lifespan.  The model reproduced classic predictions that social
learning can yield higher payoffs than individual learning when uncertainty is
sufficiently limited. We found increased
payoff ambiguity and lifespan led to greater drift, which is essentially
evolutionary uncertainty in a finite-size population. 
Our model helped us understand that social
learning acts as a scaffold for individual learning---they are not opposites. 
We endowed agents
with an evolutionarily- and biologically-plausible individual learning
mechanism, softmax search, that provided strong individual
learning that helped agents recover
from receiving outdated information. This enabled the evolution of social learning 
even when social information is likely outdated. 
Detailed theoretical
models like this are critical for predicting how humans might adapt to an 
uncertain and rapidly changing world.
This knowledge could help suggest plausibly beneficial interventions to mitigate 
existential threats~\cite{Moya2020,Jones2021}, especially in
the most vulnerable communities~\cite{McNamara2020}, and to capitalize on new
behavioral opportunities such as transitioning to clean energy use and
production~\cite{NatureEnergyEditorialPromisesPremises2018,Brisbois2022}.

@@

We found that the evolution of social learning, even under only a few modest 
assumptions, is a complex phenomenon, dependent on several interrelated factors. 
By systematically
identifying and operationalizing common theoretical assumptions about contextual
uncertainty in social learning models, we developed a more systematic
understanding of which uncertainty factors interact to determine whether social
learning evolves. By explicitly modeling a key individual-level learning mechanism
shared across taxa, namely the ability to adjust behavior to uncertainty and new
information, we saw that social learning could evolve even if social learning
regularly provides outdated information. This work informs most directly our current
understanding of the evolution of social learning, but has broader significance for
understanding problem solving under uncertainty. Our model and its software
implementation were designed to be modular and extensible.  Indeed, we plan to extend
this model to further develop a more detailed theoretical understanding of the
evolution of social learning; we hope others will, too.

@@

Model populations reached fixation within few to 140 generations on
average (Figure~\ref{fig:steps}).
Note 140 generations is equivalent to 2800 time steps. We set the maximum 
number of iterations globally to $20~000$, which is 1000 generations when
$L=20$. Although we can be relatively sure that this maximum number of 
generations is rarely reached based on our main text analysis of $\meanG$,
we additionally tabulated the number of simulations that did not converge to
fixation
within this step limit. We found that 0.1\% of simulations did not converge
when $B=2$, 0.2\% did not converge when $B=4$, and 2.0\% did not converge
when $B=10$ (Table~\ref{tab:convergence}). When a trial did not converge we
still included that trial's $\meansl$ value in the average across trials,
but for these $0 < \meansl < 1$.

@@

\subsection{Number of prospective teachers sensitivty analysis}
